### 2021-02-27 [2020-08]_ACMTOMS_(MQSI)/paper/4-analysis.tex
Next round of response done, sending to coauthors.

\heading{4. COMPLEXITY AND SENSITIVITY}
Algorithms 1, 2, and 4 have ${\cal O}(1)$ runtime. Given a fixed
schedule for shrinking derivative values, Algorithm 3 has a ${\cal
  O}(n)$ runtime for $n$ data points. In execution, the majority of
the time, still ${\cal O}(n)$, is spent solving the banded linear
system of equations for the B-spline coefficients. Thus for $n$ data
points, the overall execution time is ${\cal O}(n)$. The quadratic
facet model produces a unique sensitivity to input perturbation, as
small changes in input may cause different quadratic facets to be
associated with a breakpoint, and thus different initial derivative
estimates can be produced. This phenomenon is depicted in Figure
1. Despite this sensitivity, the quadratic facet model is still
preferred because it generally provides aesthetically pleasing (low
{\it wiggle}) initial estimates of the first and second derivatives at
the breakpoints while perfectly capturing local low-order phenomena
(e.g., a run of points on a straight line) in the data. The binary
search for a point on the monotone boundary in $(\tau_1$, $\alpha$,
$\beta$, $\gamma)$ space is preferred because it results in monotone
quintic spline interpolants that are nearer to initial estimates than
a policy that strictly shrinks derivative values.

>>>>

\heading{4. COMPLEXITY AND SENSITIVITY}
Algorithms 1, 2, and 4 have ${\cal O}(1)$ runtime. Given a fixed
schedule for shrinking derivative values, Algorithm 3 has a ${\cal
  O}(n)$ runtime for $n$ data points. In execution, the majority of
the time, still ${\cal O}(n)$, is spent solving the banded linear
system of equations for the B-spline coefficients. Thus for $n$ data
points, the overall execution time is ${\cal O}(n)$.

The quadratic facet model produces a unique sensitivity to input
perturbation, as small changes in input may cause different quadratic
facets to be associated with a breakpoint, and thus different initial
derivative estimates can be produced. This phenomenon is depicted in
Figure 1. Despite this sensitivity, the quadratic facet model is still
preferred because it exactly captures local linear and quadratic
behavior while empirically producing final approximations with less
wiggle ($L2$ norm of the second derivative) than other methods. A
weighted harmonic mean estimate of first derivatives may be more
accurate when the underlying function changes at a rate greater than a
quadratic, but that method increases the sensitivity to small
perturbations in data and empirically results in quintic splines with
greater wiggle.

The binary search for a point on the monotone boundary in $(\tau_1$,
$\alpha$, $\beta$, $\gamma)$ space is performed because it results in
monotone quintic spline interpolants with derivative values that are
absolutely nearer to initial estimates than a search that strictly
shrinks derivative values. Given the initial derivative estimates have
desirable properties (capture low-order phenomena and are low wiggle),
this search results in an approximation that is both monotone and has
derivative values similar to the initial estimates.
